---
title: "Plots for Example 1"
output: html_document
---

```{r, warning=FALSE, message=FALSE, results='hide'}
library(tidyverse)
library(cowplot)
library(latex2exp)
library(reticulate)
np = import("numpy")

example = 1

if (example == 1) {
  qlims = c(-150, 100)
} else {
  qlims = c(-100, 100)
}


rldir = "../results/rl_ex1/seed_123456"
rldf = read_csv(sprintf("%s/best_eval.csv", rldir)) %>% 
  mutate(method="RL") %>% 
  mutate(eval0 = 100 * eval0, eval1 = 100 * eval1, eval2 = 100 * eval2) %>% 
  mutate(x=50*x)

bsddir = "../results/bsd_ex1/"
bsddf = read_csv(sprintf("%s/uhatdata.csv", bsddir))  %>% 
  rename(x=t, y=pt) %>% 
  mutate(method="BSD")
```

Plot the RL state-action value function for Example 1.

```{r, fig.height=4, fig.width=11}
brks = c(-150, 0, 100)

df = rldf %>% 
  dplyr::select(x, y, eval0, eval1, eval2) %>%
  pivot_longer(cols=c("eval0", "eval1", "eval2"), names_to="D", values_to="val") %>%
  mutate(D=as.integer(str_replace(D, "eval", "")))
# head(df)

levs = c(
  "0"=TeX("$D_t=0$"),
  "1"=TeX("$D_t=1$"),
  "2"=TeX("$D_t=2$")
)
df$D_ = as.factor(df$D)
levels(df$D_) = levs

p = df %>% 
  mutate(val = pmin(pmax(val, qlims[1]), qlims[2])) %>% 
  ggplot() +
  geom_tile(aes(x=x, y=y, fill=val)) +
  facet_wrap(~ D_, ncol=3, labeller = label_parsed) +
  theme_cowplot() +
  scale_fill_gradientn(
    limits=qlims,
    values = scales::rescale(brks),
    colors=c("black", "grey", grey(0.9))
  ) +
  labs(fill="Value", x=TeX("Time step $\\t$"), y=TeX("Running mean ($\\p_t$)")) +
  guides(fill = guide_colourbar(barwidth = 20, title.vjust=1)) +
  theme(
    strip.background = element_rect(fill="white"),
    strip.text = element_text(size=15, vjust=-1),
    legend.position = "top",
    legend.margin=margin(0,0,0,0),
    legend.title.align = 0.5,
    legend.box.margin=margin(0,-5,-20,-5),
    strip.text.x=element_text(vjust=-1)
  )

# plot_grid(p, labels="B)", vjust=3)
ggsave(filename="paper_figures/ex1_qvals.png", plot=p, width=10, height=4.5, units="in")
p
```

Plot the fitted utility function from constrained backward induction in Example 1 to compare with the above.

```{r, fig.height=4, fig.width=11}
brks = c(-150, 0, 100)

df = bsddf %>% 
  dplyr::select(x, y, uhat0, uhat1, uhat2) %>%
  pivot_longer(cols=c("uhat0", "uhat1", "uhat2"), names_to="D", values_to="val") %>%
  mutate(D=as.integer(str_replace(D, "uhat", "")))

levs = c(
  "0"=TeX("$D_t=0$"),
  "1"=TeX("$D_t=1$"),
  "2"=TeX("$D_t=2$")
)
df$D_ = as.factor(df$D)
levels(df$D_) = levs

p = df %>% 
  mutate(val = pmin(pmax(val, brks[1]), brks[3])) %>% 
  ggplot() +
  geom_tile(aes(x=x, y=y, fill=val)) +
  facet_wrap(~ D_, ncol=3, labeller = label_parsed) +
  theme_cowplot() +
  scale_fill_gradientn(
    limits=qlims,
    colors=c("black", "grey", grey(0.9)),
    values=scales::rescale(brks)
  ) +
  labs(fill="Utility", x=TeX("Time step $\\t$"), y=TeX("Running mean ($\\p_t$)")) +
  guides(fill = guide_colourbar(barwidth = 20, title.vjust = 1)) +
  theme(
    strip.background = element_rect(fill="white"),
    strip.text = element_text(size=15, vjust=-1),
    legend.position = "top",
    legend.margin=margin(0,0,0,0),
    legend.title.align = 0.5,
    legend.box.margin=margin(0,-5,-20,-5),
    strip.text.x=element_text(vjust=-1)
  )

# plot_grid(p, labels="A)", vjust=3)
ggsave(plot=p, filename="paper_figures/ex1_uhat.png", width=10, height=4.5, units="in")
p
```


Plot the decision boundaries in Example 1.

```{r}
tracedata = read_csv(paste0(bsddir, "tracedata.csv"))
```

Plot RL decision boundaries.

```{r}
# No segments
p = ggplot(rldf) +
  geom_tile(aes(x=x, y=y, fill=factor(act))) + 
  scale_fill_manual(values=c(grey(0.95), "grey", "black")) +
  geom_segment(
    aes(x=x, y=y, xend=xend, yend=yend, color=as.factor(num)),
    arrow=arrow(length=unit(0.12,"cm")),
    alpha=0.6,
    size=0.8,
    data=na.omit(tracedata)
  ) +
  guides(color=FALSE) +
  scale_color_hue(l=40, c=40) +
  theme_cowplot() +
  theme(
    legend.position="top",
    legend.margin=margin(0,0,0,0)
    ) +
  labs(fill="Decisions", x=TeX("Time step $t$"), y=TeX("Running mean ($p_t$)"))
ggsave("paper_figures/ex1_qvals_boundaries.png", plot=p, height=4, width=5, units="in")
p
```


Plot BSD decision boundaries.

```{r}
p = ggplot(bsddf) +
  geom_tile(aes(x=x, y=y, fill=factor(dstar))) + 
  scale_fill_manual(values=c(grey(0.95), "grey", "black")) +
  geom_segment(
    aes(x=x, y=y, xend=xend, yend=yend, color=as.factor(num)),
    arrow=arrow(length=unit(0.12,"cm")),
    alpha=0.6,
    size=0.8,
    data=na.omit(tracedata)
  ) +
  guides(color=FALSE) +
  theme_cowplot() +
  theme(
    legend.position="top",
    legend.margin=margin(0,0,0,0)
    ) +
  labs(fill="Decisions", x=TeX("Time step $t$"), y=TeX("Running mean ($p_t$)")) +
  scale_color_hue(l=40, c=40)

ggsave("paper_figures/ex1_uhat_boundaries.png", plot=p, height=4, width=5, units="in")
p
```

Historical rewards from Q-learning training (may vary greatly from seed to seed).

```{r}
evals = np$load(
  sprintf("%s/evaluations.npz", rldir), allow_pickle = TRUE
)
results = 100 * evals$get("results")
evaldf = tibble(
  timesteps = evals$get("timesteps"),
  mu = apply(results, 1, mean),
  upper = apply(results, 1, quantile, 0.95),
  lower = apply(results, 1, quantile, 0.05)
)
maxr = max(evaldf$mu)

p = ggplot(evaldf) +
  geom_line(aes(x=timesteps, y=mu, color="Mean of 1000 test episodes")) +
  geom_hline(aes(yintercept=maxr, lty="Best eval score")) +
  theme_cowplot() +
  labs(x="Simulation steps", y=TeX("Evaluation episode returns (G_0)")) + 
  scale_fill_manual(values="black") +
  scale_color_manual(values="black") + 
  scale_linetype_manual(values=2) + 
  theme(
    legend.position="top",
    legend.title=element_blank(),
    legend.spacing.y = unit(0, "cm"),
    legend.margin=margin(0,0,0,0)
  ) +
  labs(fill="", color="") + 
  scale_y_continuous(n.breaks=10)

ggsave("paper_figures/ex1_returns.png", plot=p, width=5, height=4, units="in")
p
```

The maximum mean reward over 1000 eval episodes is

```{r}
maxr
```
